---
title: "Biol 3872 Lab3-Assignment"
output: html_document
date: "`r Sys.Date()`"
author:
student_id: ""
editor_options: 
  chunk_output_type: console
---

**BEFORE STARTING:**\
1. Include your name (as shown in Brightspace) in the `author` section above as text (e.g., "Liam MacNeil").\
2. Install and load the packages required to run your code chunks/

Assignments will largely be evaluated based on your code chunk answers. If they are incomplete or formatted incorrectly (e.g., code chunks don't run), no marks will be awarded. I recommend experimenting with solving these answers in a separate script/markdown document and reporting your answers with minimal script.

```{r}
library(tidyverse)
data("iris")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

::: {style="background-color: #44AA99; padding: padding: 5px; border-radius: 2px;"} \
## 1. Task 1

A)  What species displays the longest petal lengths? Store the species name as a character called `longest`.
B)  What species has the greatest variance in sepal width? Store the species name as a character called `variable`.

Report your answer in the code chunks below.

```{r Task 1}
#A) 
# YOUR CODE HERE

#B)
# YOUR CODE HERE

```
:::
\

::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 2. Task 2

A)  Using `base` R, produce a `plot` of all variables excluding the `species` variable but color points by species identity. Save this plot as a .png called `base_plot.png` in your `Figures` folder. Specify figure dimensions as 12 x 12 cm with 300 DPI.

```{r Task 2}
#A)
# YOUR CODE HERE

```
:::

\

::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 3. Task 3

A)  Based on the grammar of graphics, produce a **blank plot (no data plotted)** named `blank` using the `iris` data set with species as the x axis and sepal widths as the y axis. Store this object as a `ggplot` object named `empty_plot`.

```{r Task 3}
#A)
# YOUR CODE HERE
```
:::

\

::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 4. Task 4

A)  Fix the code below so that it runs and plots successfully and store it as a plot object named `fixed_traits_plot`. This plot should have petal length as the x-axis, sepal length as y axis and point size should correspond to petal width. Each point should be colored by species name.

```{r Task 4}
# FIX AND IMPLEMENT CODE BELOW
#ggplot(iris, aes(x = Petal.Length, y = Sepal.Lenth), color = "Species")+
 # geom_line(size = Petal.Width)+
  #theme_minimal()

# A)
# YOUR CODE HERE
```
:::

\

::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 5. Task 5

A)  Fix and modify the code below and add a discretized `geom` (e.g., `hex`) with exactly 10 bins, with a `viridis` filled palette (check `?` for different `scale_fill_...` options), and facets for each species. Name and store it as a plot object as `fixed_traits_plot_discrete`. *Clue: Can you fill a hexagon with species names using a continuous palette?*
```{r Task 5}
# FIX AND IMPLEMENT CODE BELOW

#A) 
#ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species))+
#  geom_line(alpha = 0.5, aes(as.factor(Petal.Length)))+
#  theme(legend.position = "left")+
#  scale_x_continuous(limits = seq(0,Inf,2))

# YOUR CODE HERE

```
:::


::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 6. Task 6

A)  Using the hotkey to generate a pipe operator in the `tidyverse`, pivot the petal measurements in the `iris` dataset to a longer format and successfully generate a boxplot with these petal measurements as the y-axis and measurement type (e.g., length v. width) as the x axis.Use these unique color codes (`"forestgreen"`,`"skyblue"`,`"maroon"`) to **fill**- not color- each species. This should be generated in a single code chunk. Add the `minimal` theme with text size at 12. 

Name and store it as a plot object as `petal_plot`.
```{r Task 6}
# YOUR CODE HERE

#A)
# YOUR CODE HERE

```
:::

::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 7. Task 7

A) Similar to `Task 6`, in a single code chunk, pivot the petal measurements in the `iris` dataset to a longer format. Generate a scatter plot comparing with different measurement types on the x axis and their values on the y axis. Add a jitter `geom` and give each species a unique `facet` with a free (unique) y-axis. Name and store it as a plot object as `facet_petal_plot`.

B)  Reproduce this same plot named `facet_petal_plot_reorder` where the facets are reversed by the reordering the `species` factor.
```{r Task 7}
# YOUR CODE HERE

#A)
# YOUR CODE HERE

#B)
# YOUR CODE HERE
```
:::

::: {style="background-color: #44AA99; transparent; padding: 5px; border-radius: 2px;"}

## 8. Task 8

A) Create a new column in the `iris` dataset with `base` R or `dplyr` called `petal_area` which compues petal area.

B)  Create and store a ggplot object named `area_plot` comparing `petal_area` (x axis) to `Sepal.Length` (y axis) with points colored by species.

```{r Task 8}
#A)
# YOUR CODE HERE
#B)
# YOUR CODE HERE
```
:::

## 8. Task 8C

C) Save the previous plot as a `ggplot` object in your Figures folder named `area_plot.png` and define dimensions as 12x12 cm with 300 DPI.

```{r}
#C)
# YOUR CODE HERE
```
Be sure before submitting this assignment to clear your and environment and re-run your code to ensure it works then submit this `markdown` file to the relevant Brightspace submission box.
